---
title: "homework7"
author: "Nick Maurizi"
date: "2025-03-06"
output: html_document
---

1. Getting packages and data. iris is a data frame with 150 rows and five variables (columns).
```{r}
library(tidyverse)
data(iris)
```

2. Creating iris 1 by filtering for only rows that meet the desired criteria.
```{r}
iris1 <- filter(iris, Species%in% c("virginica","versicolor"),Sepal.Width > 2.5,Sepal.Length>6) 
```

3. Creating iris 2 that only includes desired columns using the select function
```{r}
iris2 <- select(iris1, Species, Sepal.Length, Sepal.Width)
```

4. Orders the data by descending sepal length.
```{r}
iris3 <- arrange(iris2, by=desc(Sepal.Length))
```

5. Adding a sepal area (L x W) to the previous data frame. Uses mutate to provide the formula and create the new column simultaneously.
```{r}
iris4 <- mutate(iris3, Sepal.Area=Sepal.Length*Sepal.Width)
```

6. Creates a small data frame that calculates averages.
```{r}
iris5 <- summarize(iris4, average.sepal.length= mean(Sepal.Length), average.sepal.width=mean(Sepal.Width),sample.size=n())
print(iris5)
```

7. Now we do averages by species.
```{r}
iris6 <- group_by(iris4, Species)
iris6 <- summarize(iris6, average.sepal.length= mean(Sepal.Length), average.sepal.width=mean(Sepal.Width),sample.size=n())
print(iris6)
```

8. This chunk shows how using pipe statements can speed things up.
```{r}
irisFinal <- iris%>% 
  filter(Species%in% c("virginica","versicolor"),Sepal.Width > 2.5,Sepal.Length>6)%>% 
  select(Species, Sepal.Length, Sepal.Width)%>%
  arrange(by=desc(Sepal.Length))%>%
  mutate(Sepal.Area=Sepal.Length*Sepal.Width)%>%
  group_by(Species)%>%
  summarize(average.sepal.length= mean(Sepal.Length), average.sepal.width=mean(Sepal.Width),sample.size=n())

irisFinal==iris6  #checking my work
```

9. Formating the original iris data frame "the long way" using pivot functions.
```{r}
irislonger <- iris%>%
  select(Species,Sepal.Length,Sepal.Width,Petal.Length,Petal.Width)%>%
  pivot_longer(cols = 2:5, names_to = "Measure", values_to = "Value")
```
