---
title: "homework4"
author: "Nick Maurizi"
date: "2025-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. I began by assigning the values to their variables, then evaluating each expression and assigning to z, and displaying z

```{r}

x <- 1.1
a <- 2.2
b <- 3.3

#Expressions
z <- x^(a^b)
print(z)

z <- (x^a)^b
print(z)

z <- (3*x^3+2*x^2+1)
print(z)


```

2. 
```{r}
vector_a <- c(1:8,7:1)
print(vector_a)

temp_vec <- c(1:5)
vector_b <- rep(temp_vec,times=temp_vec)
print(vector_b)

temp_vec2 <- c(5:1)
vector_c <- rep(temp_vec2,times=temp_vec)
print(vector_c)


```

3. I assigned the variables their values by picking them out of the vector I created. I found r using the Pythagorean theorem and then used inverse sin to find theta. 
```{r}
vector <- runif(2)

x <- vector[1]
print(x)
y <- vector[2]
print(y)

r <- (x^2 + y^2)^(1/2) 
print(r)

theta <- asin(y/r)
print(theta)



```

4. Noahs arc
```{r}
queue <- c("sheep", "fox", "owl", "ant") 
print(queue)
queue <- c(queue,"serpent")
print(queue)
queue <- queue[2:5]
print(queue)
queue <- c("donkey",queue)
print(queue)
queue <- queue[1:4]
print(queue)
queue <- queue[c(1,2,4)]
print(queue)
queue <- c(queue[1:2],"aphid",queue[3])
print(queue)
which(queue=="aphid")


```

5. I used the subset function here to include all numbers that cannot be divided evenly (defined as having a remainder not equal to 0) by 2, 3, or 7.
```{r}

math_vector <- c(1:100)

math_final <- subset(math_vector, (math_vector%%2 != 0) & (math_vector%%3 !=0) & (math_vector%%7 != 0))
print(math_final)


```















